Wealth dynamics in the presence of network structure and primitive cooperation

### 2019-06-27 manuscript/draftcssa19.tex
section 1 completed

The current work originated in our attempts to incorporate and tease out the effects of social structures in simple models of wealth dynamics in the presence of environmental stochasticity, and a simple form of resource pooling. These models were inspired by our search for analogs of the grid world agent-based models (ABMs) of Friesen and Mudigonda \footnote{Include proper citation of SriMil here} where foraging agents that pool their resources were shown to on-average outperform non-pooling agents. The original model used foraging to mix the population and create opportunities for interaction, and, when certain conditions were met, allowed resource sharing. Our model achieves agent interaction by postulating a static network structure which partially mixes the agents.

@@

Models used in scientific inquiry serve a specific purpose. In this work, the goal is to construct simple toy models that reproduce non-trivial wealth inequality distributions in the presence and absence of primitive forms of cooperation, clearly delineating the role of network structure in generating wealth inequality. We make no suggestions that these models \textit{explain} the phenomena of interest; we are only interested in constructing the simplest possible models, with no detailed empirical grounding, but with the potential to generate realistic looking inequality distributions. In doing so, we aspire to shed light on the \textit{true} social and economic mechanisms underlying the genesis and persistence of economic inequality.

%Briefly explain the model here and then move on to the next section. .

%The current plan is start with economics economic sociology and wealth dynamic models. The former is discussed in Jackson's review, the latter in rvenkat's CSSA18 paper. Then we move on to how wealth dynamics and inequality questions matter also to other fields like anthropology and sociology. Wealh dynamics in such situations can also be addressed using economic models and is the focus of our manuscript.

@@

\section{Model}
~\cite{redner2001guide}
%Use chapter 3 of redner's book here. 
%Also, the network model presented here seem to be one of a kind. Typically dynam ics on networks are deterministic in nature. The models discussed here is different from deterministic dynamic processes and epidemic processes, and diffusion type processes (percolation processes). The models are stochastic dynamic systems on networks. Such social processes are studied using Stochastic Actor Oriented Models (SOAM) Snidjers. ALso use Kolacyzk and Csardi's discussion as reference. Social influence models..... and others) This is not directly about influence but more about cooperation. Need to find a nice way of talking about this.

>>>>

The current work originated in our attempts to incorporate and tease out the effects of social structures in simple models of wealth dynamics in the presence of environmental stochasticity, and a simple form of resource pooling. These models were inspired by our search for analogs of the grid world agent-based models (ABMs) of Friesen and Mudigonda~\cite{srimil} where foraging agents that pool their resources were shown to on-average outperform non-pooling agents. The original model used foraging to mix the population and create opportunities for interaction, and, when certain conditions were met, allowed resource sharing. Our model achieves agent interaction by postulating a static network structure which partially mixes the agents.

@@

Models constructed here, like those used in scientific inquiry in general, serve a specific purpose. In this work, we construct simple toy models to reproduce certain aspects of non-trivial wealth inequality distributions in the presence and absence of primitive forms of cooperation, clearly delineating the role of network structure in generating wealth inequality. We make no suggestions that these models \textit{explain} the phenomena of interest; we are only interested in constructing the simplest possible models, with no detailed empirical grounding, but with the potential to generate realistic looking inequality distributions. In doing so, we aspire to shed light on the \textit{true} social and economic mechanisms underlying the genesis and persistence of economic inequality.

%What is the scientific question of interest to us? 
%How features of models of structure and dyanmics intera

The phenomenon under scrutiny is the emergence and evolution of economic inequality in societies with non-trivial social structures. The model we use to answer questions surrounding this phenomena must possess a dynamic model of wealth accumulation, a model of social structure; and a suitably useful measure of inequality. The dynamics are modeled as Brownian noise driven linear dynamical systems, here constant growth rate; the structure is modeled by networks, here random graphs; the measure of inequality used is Gini coefficients. Although both the dynamical system and the network model are quiet well understood, the precise interplay of structure and dynamics produces interesting emergent wealth distributions. To the best of our knowledge, this specific combination of network structure and social dynamics has not been discussed in the computational social science or mathematical sociology literature and consider this the primary contribution of this work.

Apart from enabling interactions between social actors, social structures like institutions also shape the form of cooperation and coordination mechanisms. The institutions can take the form of economic institutions, like banks and cooperatives; or the form of norms, like resource sharing practices in societies. The original SriMil model~\cite{srimil}consists of simple resource pooling arrangement where aggregates of agents pool their excess wealth in a common institution called protoinstitutions, agreeing to provide this saved resource to individual agents in times of need. The model of resource pooling used in this work is identical to the SriMil model.

While the analysis to be presented in later sections focus only on homogeneous agents with simple drift-diffusion dynamics on Erd\H{o}s-Renyi network models, space constraints prevent us from repeating the analysis on other standard \textit{textbook} networks like scale free and small world networks. We also address more empirically grounded models in the concluding section of this paper.

In the next section, we discuss the mathematical formulation of the model. In section 3, we present the analysis of our simulation experiments, summarize key findings and discuss why we chose \texttt{julia} programming language~\cite{Julia-2017} for our implementation. In section 4, we discuss the limitations of our simple models, extensions to dynamics and networks more expressive than the ones presented here, and planned future work.

@@

\section{Model}
~\cite{redner2001guide} ~\cite{rv}
%Use chapter 3 of redner's book here. 
%Also, the network model presented here seem to be one of a kind. Typically dynam ics on networks are deterministic in nature. The models discussed here is different from deterministic dynamic processes and epidemic processes, and diffusion type processes (percolation processes). The models are stochastic dynamic systems on networks. Such social processes are studied using Stochastic Actor Oriented Models (SOAM) Snidjers. ALso use Kolacyzk and Csardi's discussion as reference. Social influence models..... and others) This is not directly about influence but more about cooperation. Need to find a nice way of talking about this.
